Reproduce anaylses of Beninca et al (2008)

### 2016-02-19 code/report.Rmd
checking global LE

Read in the data and remove some blank rows and columns.
```{r}
spp.abund <- read.csv(text=getURL(paste0(repo,"/data/species_abundances_original.csv")), skip=7, header=T)

@@

Bring in the nutrient data:
```{r}
nuts <- read.csv(text=getURL(paste0(repo,"/data/nutrients_original.csv")), skip=7, header=T)
#nuts <- read.csv("~/Dropbox (Dept of Geography)/RREEBES/Beninca_etal_2008_Nature/data/nutrients_original.csv", skip=7, header=T)

@@

The exact method for identifying the "long sequences of zeros" was not given. However, the transformed data with zeros removed was provided, and from this we can get the day range of the retained data:
```{r}
tr <- read.csv(text=getURL(paste0(repo,"/data/transformed_data_Nature2008.csv")),
               skip=7, na.string="")
tr <- tr[,-14:-24] ## remove bad columns
tr <- tr[-693:-694,] ## remove last two rows (contain summary stats)
tr <- gather(tr, key="Species", value="Abundance", 2:13) ## make long format
levels(tr$Species)[levels(tr$Species)=="Calanoids"] <- "Calanoid.copepods" # rename a level (and next line)
levels(tr$Species)[levels(tr$Species)=="Total.Dissolved.Inorganic.Nitrogen"] <- "Total.dissolved.inorganic.nitrogen"
names(tr)[2] <- "variable" # rename a couple of variables...
names(tr)[1] <- "Day.number"
tr$Day.number <- as.numeric(as.character(tr$Day.number)) # make Day.number numeric
## get day range, in order to use this to constrain later data
min.max.day <- group_by(tr, variable) %>%
  summarise(min.day=min(Day.number[!is.na(Abundance)]), max.day=max(Day.number[!is.na(Abundance)]))
```

@@

Compare correlations in table 1 of the original article with those calculated here:
```{r}
# read in the correlations given in the original publication
original.cors <- read.csv(text=getURL(paste0(repo,"/data/table1_original_article.csv")), skip=0, header=T, row.names = 1)
original.cors <- as.matrix(original.cors)

@@

Compare graphically with original LEs
```{r}
## read in LE values given in original publication
original.LEs <- read.csv(text=getURL(paste0(repo,"/data/original_direct_LEs.csv")), skip=0, header=T)
both_LEs <- full_join(original.LEs, LEs) # merge the reproduced and original values
ggplot(both_LEs, aes(x=LE, y=le)) +
  geom_point() +
  xlim(0.02, 0.09) +
  ylim(0.02, 0.09) +
  geom_abline(intercept=0, slope=1, colour="red") +
  xlab("Lyapunov exponent from original article") + 
  ylab("Reproduced Lyapunov exponent")
```

@@

Check this against the data direct from Steve:
```{r}
from.steve <- read.csv(text=getURL(paste0(repo,"/data/interp_short_allsystem_newnames.csv")), header=T)
from.steve <- gather(from.steve, Species, Abundance, 2:13)
names(from.steve) <- c("Day.number", "variable", "value")
from.steve$value_steve <- from.steve$value 
from.steve <- select(from.steve, -value)

@@

## joint the data, again first making days character, then back to numeric after
mt$Day.number <- as.character(mt$Day.number)
from.steve$Day.number <- as.character(from.steve$Day.number)
ff <- inner_join(mt, from.steve)
mt$Day.number <- as.numeric(mt$Day.number)
from.steve$Day.number <- as.numeric(from.steve$Day.number)
ff$Day.number <- as.numeric(ff$Day.number)

@@

@@

@@

## load the already saved data from github (this can take some time depending on the internet connection)
source_data(paste0(repo, "/data/GLE_estimate.Rdata?raw=True"))

@@

## read in the data of the original article that came direct from Elisa
original_preds <- read.csv(text=getURL(paste0(repo,"/data/original_rsquared.csv")), skip=0, header=T)
opw <- gather(original_preds, Species, r_squared, 2:25)
opw <- separate(opw, Species, c("Species", "Model"))

>>>>

Read in the data and remove some blank rows and columns.
```{r}
spp.abund <- read.csv(text=getURL(paste0(repo,"/data/original/species_abundances_original.csv")), skip=7, header=T)

@@

Bring in the nutrient data:
```{r}
nuts <- read.csv(text=getURL(paste0(repo,"/data/original/nutrients_original.csv")), skip=7, header=T)

@@

The exact method for identifying the "long sequences of zeros" was not given. However, the transformed data with zeros removed was provided, and from this we can get the day range of the retained data:
```{r}
tr <- read.csv(text=getURL(paste0(repo,"/data/original/transformed_data_Nature2008.csv")),
               skip=7, na.string="")
tr <- tr[,-14:-24] ## remove bad columns
tr <- tr[-693:-694,] ## remove last two rows (contain summary stats)
tr <- gather(tr, key="Species", value="Abundance", 2:13) ## make long format
levels(tr$Species)[levels(tr$Species)=="Calanoids"] <- "Calanoid.copepods" # rename a level (and next line)
levels(tr$Species)[levels(tr$Species)=="Total.Dissolved.Inorganic.Nitrogen"] <- "Total.dissolved.inorganic.nitrogen"
names(tr)[2] <- "variable" # rename a couple of variables...
names(tr)[1] <- "Day.number"
tr$Day.number <- as.numeric(as.character(tr$Day.number)) # make Day.number numeric
## get day range, in order to use this to constrain later data
min.max.day <- group_by(tr, variable) %>%
  summarise(min.day=min(Day.number[!is.na(Abundance)]), max.day=max(Day.number[!is.na(Abundance)]))
```

@@

Compare correlations in table 1 of the original article with those calculated here:
```{r}
# read in the correlations given in the original publication
original.cors <- read.csv(text=getURL(paste0(repo,"/data/original/table1_original_article.csv")), skip=0, header=T, row.names = 1)
original.cors <- as.matrix(original.cors)

@@

Compare graphically with original LEs
```{r}
## read in LE values given in original publication
original.LEs <- read.csv(text=getURL(paste0(repo,"/data/original/original_direct_LEs.csv")), skip=0, header=T)
both_LEs <- full_join(original.LEs, LEs) # merge the reproduced and original values
ggplot(both_LEs, aes(x=LE, y=le)) +
  geom_point() +
  xlim(0.02, 0.09) +
  ylim(0.02, 0.09) +
  geom_abline(intercept=0, slope=1, colour="red") +
  xlab("Lyapunov exponent from original article") + 
  ylab("Reproduced Lyapunov exponent")
```

@@

Check this against the data direct from Steve:
```{r}
from.steve <- read.csv(text=getURL(paste0(repo,"/data/original/interp_short_allsystem_newnames.csv")), header=T)
from.steve <- gather(from.steve, Species, Abundance, 2:13)
names(from.steve) <- c("Day.number", "variable", "value")
from.steve$value_steve <- from.steve$value 
from.steve <- select(from.steve, -value)

@@

## join the data, again first making days character, then back to numeric after
mt$Day.number <- as.character(mt$Day.number)
from.steve$Day.number <- as.character(from.steve$Day.number)
ff <- inner_join(mt, from.steve)
mt$Day.number <- as.numeric(mt$Day.number)
from.steve$Day.number <- as.numeric(from.steve$Day.number)
ff$Day.number <- as.numeric(ff$Day.number)

@@

```{r, eval=F, echo=F}
## code to export the interpolated data, with which to test the GAM functions
mt_wide <- spread(mt, key=variable, value=value)
write.csv(mt_wide, "~/Desktop/Repro_data_for_gam_test.csv")
```

@@

```{r, eval=F, echo=F}
## code to export the interpolated and transformeddata, with which to test the GAM functions
write.csv(X, "~/Desktop/Repro_data_for_gam_test_transformed.csv")
```

@@

## load the already saved data from github (this can take some time depending on the internet connection)
source_data(paste0(repo, "/data/reproduction/GLE_estimate.Rdata?raw=True"))

@@

## read in the data of the original article that came direct from Elisa
original_preds <- read.csv(text=getURL(paste0(repo,"/data/original/original_rsquared.csv")), skip=0, header=T)
opw <- gather(original_preds, Species, r_squared, 2:25)
opw <- separate(opw, Species, c("Species", "Model"))
